---
title: "COP22 FAST Datapack Code"
author: "Jairo Montes & Vanessa Da Costa"
date: "04/13/2023" #LAST UPDATE
output: html_document
---


##Purpose:
This code was developed to integrate real-time notional budget data from the COP22 FAST with COP22 notional target data from  Data Pack and historic expenditure and performance data. 


##Step 1: Installing Packages
R packages are a collection of R functions, complied code and sample data. They are stored under a directory called "library" in the R environment. By default, R installs a set of packages during installation. https://www.tutorialspoint.com/r/r_packages.htm
```{r}
install.packages("tidyr")
install.packages("tidyverse")
install.packages("dplyr")
install.packages("stringr")
install.packages("readxl")
install.packages("here")
#install.packages("googlesheets4")
install.packages("data.table")
install.packages("splitstackshape")
#install.packages("googledrive")
devtools::install_github("USAID-OHA-SI/glamr")
install.packages("purrr")
install.packages("devtools")
install.packages("remotes")
remotes::install_github("USAID-OHA-SI/tameDP")
remotes::install_github("USAID-OHA-SI/gophr")
install.packages("keyring")
```

##Step 2: Running Libraries
```{r}
library(rmarkdown)
library(stringr)
library(dplyr)
library(tidyverse)
library(tidyr)
library(readr)
library(readxl)
library(glamr)
library(purrr)
library(data.table)
#library(googlesheets4)
library(splitstackshape)
#library(googledrive)
library(tameDP)
library(gophr)
library(devtools)
library(keyring)

library(here)

```


##Step 3: Locating the dataset files
The "here()" funcitons constructs paths to your project's files, it will locate the files relative to your project root.
```{r}
  here()
  here("C:/Users/jmontespenaloza/Documents/COP23 Tools")
  here("C:/Users/vdacosta/Documents/R/COP23 Tools/FASTS")
 
  set_email("jmontespenaloza@usaid.gov")
  glamr::load_secrets()
  set_datim("jmontespenaloza2")
  
setwd("C:/Users/jmontespenaloza/Documents")
library(here)
here()
```

##Step 4: Creating functions
Load the following source files to apply the functions necessary to create each data stream
```{R}
source("~/GitHub/EA-Utilities/COP Analytics/fast_function.R")
source("~/GitHub/EA-Utilities/COP Analytics/dp_function.R") 
#source("~/GitHub/EA-Utilities/COP Analytics/esf_function.R")
#source("~/GitHub/EA-Utilities/COP Analytics/OU_Country.csv")
```



##Step 5: READING IN DATASETS AND APPLYING THE FUNCTIONS--> dfs
Read in the FAST & Datapack and clean it using our functions. Pattern is helpful to use here but be careful that you are not picking up more than one dataset.
```{R}
#Read in FAST & DATAPACK files
#identify all the Data Pack files (pull in validated datapacks apps.datim.org/datapack)
  datapack<- list.files("C:/Users/vdacosta/Documents/R/COP23 TOOLS/DATAPACK", full.names = TRUE)

 datapack<-list.files("C:/Users/jmontespenaloza/Documents/COP23 TOOLS/DATAPACK",full.names = TRUE)


#IMPORT DATASET WITH CORRECT WHR COUNTRIES TO MECHANISM MAPPING
df.rename<- read.csv("C:/Users/vdacosta/Documents/R/COP23 TOOLS/rename.csv" , check.names = FALSE) %>% 
    rename("Mechanism ID"= "mech_code") %>% 
           dplyr::mutate(`Mechanism ID`= as.character(`Mechanism ID`))


df.rename<- read.csv("C:/Users/jmontespenaloza/Documents/COP23 TOOLS/rename.csv" , check.names = FALSE) %>% 
    rename("Mechanism ID"= "mech_code") %>% 
           dplyr::mutate(`Mechanism ID`= as.character(`Mechanism ID`))
 
#Data frame for targets
#make sure HOME tab in caribbean datapack is changed to Jamaica
#make sure HOME tab in central america datapack is changed to Brazi

df_all<-map_dfr(.x = datapack, 
                .f= ~ tame_dp(.x, type = "PSNUxIM", map_names=FALSE))
 
df_dp<- df_all %>% datapack_im_tab() %>% 
    rename("Country2"= "Country") 

df_dp<- left_join(df_dp, df.rename, by = "Mechanism ID") %>% 
  dplyr::mutate(countryname2 = case_when(`Mechanism ID`== "86252" | `Mechanism ID`== "100129"| `Mechanism ID`== "100139"|  `Mechanism ID`== "81114"|  `Mechanism ID`== "100110"|  `Mechanism ID`== "100105"|  `Mechanism ID`== "84311" |  `Mechanism ID`== "100130"|  `Mechanism ID`== "100140"|  `Mechanism ID`== "81144"|  `Mechanism ID`== "100106"|  `Mechanism ID`== "100111"|  `Mechanism ID`== "84230" |  `Mechanism ID`== "100131"|  `Mechanism ID`== "100141"|  `Mechanism ID`== "81176"|  `Mechanism ID`== "100112"|  `Mechanism ID`== "84360" |  `Mechanism ID`== "81325"|  `Mechanism ID`== "100113"| `Mechanism ID`== "100138"|   `Mechanism ID`== "100143"| `Mechanism ID`== "100114"|  `Mechanism ID`== "84730"  ~ Country, TRUE ~ Country2)) %>% 
    select (-c(Country2, Country)) %>% 
    rename("Country"= "countryname2") 

#identify all the FAST files  
  Fast<- list.files("C:/Users/jmontespenaloza/Documents/COP23 TOOLS/FASTS WHR", full.names = TRUE)
  #RUN FIRST with step 6
  Fast<- list.files("C:/Users/vdacosta/Documents/R/COP23 TOOLS/FASTS", full.names = TRUE)
  #RUN SECOND with step 6.2
  Fast<- list.files("C:/Users/vdacosta/Documents/R/COP23 TOOLS/FASTS WHR", full.names = TRUE)

  #Data frame for Intervention Budget
  df.Intervention <- purrr::map_dfr(.x = Fast,
                                  .f = ~ FAST_Intervention(.x))

  #Data frame for Cross-Cutting Attribution Budget
  df.CCA <- purrr::map_dfr(.x = Fast,
                         .f = ~ FAST_CCA(.x))

  #Data frame for Initiative Budget
  df.Initiative <- purrr::map_dfr(.x = Fast,
                           .f = ~ FAST_Initiative(.x))

  #additional steps for Commodities
  #Data frame for Commodities Budget (prior to merging in identifiers)
  df.Commodities_pre <- purrr::map_dfr(.x = Fast,
                                  .f = ~ FAST_Commodities(.x))

  #Data frame for Mech List
  df.MechsList <- purrr::map_dfr(.x = Fast,
                                       .f = ~ FAST_MECHSLIST(.x))
  
  #Merge in  Mech List to make final Commodities data frame
  #Note that this includes OU but not Operating Unit
  #df.Commodities<- left_join(df.Commodities_pre, df.MechsList, by = "Mechanism ID")
  
  #Data frame for Earmarks from FAST by IM
  df.Earmarks_IM <- purrr::map_dfr(.x = Fast,
                                 .f = ~ FAST_Earmarks_IM(.x))

  
  
#Data frame for Operating Unit and Country
 
   df.OU_Country <- read.csv (
    "~/GitHub/EA-Utilities/COP Analytics/OU_Country.csv", check.names = FALSE,  
    fileEncoding="UTF-8-BOM")


```
##Step 6: Bind Dataframes
Master COP22 FAST/DATAPACK File
Note: Bind dataframes do not include df. Commodities_pre but only include 'df.Commodities'
```{r}
  #bind all data frames
   df.COP22PlanningDataset<-bind_rows(df.Intervention,df.CCA, df.Initiative,  
                                    df.Commodities_pre, df.Earmarks_IM) %>% 
  COP22_master_clean()
  
  df.COP22PlanningDataset<-left_join(df.COP22PlanningDataset, df.MechsList, by = "Mechanism ID")     %>% dplyr::rename("Country" = `Operating Unit`)
      
  
  #add in Operating Unit
   df.COP22PlanningDataset <- merge(df.COP22PlanningDataset, df.OU_Country, by = "Country")
  #RUN IF HAVE FASTS AND DATAPACK 
  df.COP22PlanningDataset <-bind_rows(df.COP22PlanningDataset,df_dp) 


```
##Step 6.2: Bind Dataframes WHR
```{r}
  #bind all data frames
   df.COP22PlanningDataset_WHR<-bind_rows(df.Intervention,df.CCA, df.Initiative,df.Commodities_pre, df.Earmarks_IM) %>%      
  COP22_master_clean_WHR()
  
  df.COP22PlanningDataset_WHR<-merge(df.COP22PlanningDataset_WHR, df.MechsList, by = "Mechanism ID") 
  
#final Merge 
   df.COP22PlanningDataset_Final<-bind_rows(df.COP22PlanningDataset,df.COP22PlanningDataset_WHR) %>% 
       dplyr::select(-c(`Fiscal Year`)) %>% 
     dplyr::rename("Targets" = `Target`)
     #because this is added in Tableau Prep

```
##Step 6.3: OVC Team Dataset
```{r}
  #bind all data frames
 FAST_raw<-function(df){
  df<- read_xlsx(df,"Standard COP Matrix-E", skip=3)
  
  df<-df%>% dplyr::select(-c('Funding Accounts','FAST Tabs','SCM Row ID', 'SCM ID': 'Message'))
  return(df)}

 df.raw_WHR <- purrr::map_dfr(.x = Fast,
                                  .f = ~ FAST_raw(.x))
 
 df.raw_normal <- purrr::map_dfr(.x = Fast,
                                  .f = ~ FAST_raw(.x))

 df.fast_raw <- bind_rows(df.raw_normal,df.raw_WHR)
 
```

##Step 7:Review Data
```{r}
  unique(df.COP22PlanningDataset$`Agency Category`)
  unique(df.COP22PlanningDataset_Final$`Program Area`)
  glimpse(df.Earmarks_OU)
  df.COP21PlanningDataset%>%
    group_by(`Fiscal Year`, `Data Stream`) %>%
    filter (`Fiscal Year`==2024)  %>%
    summarize (sum(`Total Planned Funding`))
```

##Step 8: Write the file to output folder.
Update the date
```{r}

## FAST + DATAPACK OUTPUT
 write.csv(df.COP22PlanningDataset_Final, paste0("C:/Users/vdacosta/Documents/R/FAST_DATAPACK", format(Sys.time(), "%d-%b-%Y"), ".csv"))

write_csv(df.COP22PlanningDataset_Final,here("C:/Users/jmontespenaloza/Documents/COP23 Tools/FAST_DATAPACK_04_13_23.csv"))


write_csv(df.fast_raw,here("C:/Users/jmontespenaloza/Documents/COP23 Tools/FAST_Raw_04_13_23.csv"))

##DATAPACK OUTPUT
  write_csv(df_all,here("C:/Users/vdacosta/Documents/R/COP22 TOOLS/Datapack_04_13_22.csv"))
```

##Additional Steps
##STEP 9: Aggregate DHI Investments Tab into its own dataset
```{r}  
    
df.DHI <-df.Intervention %>% 
      dplyr::rename("Country" = `Operating Unit`) %>% 
      dplyr::select('Country':'Digital Health Investments', 'Data Stream') %>% 
      agency_category_fast()
      distinct(.keep_all= TRUE)
      
      glimpse(df.DHI)
  
   df.DHI_Final <- merge(df.DHI, df.OU_Country, by = "Country")
  
   write_csv(df.DHI_Final,"COP22 FAST and Data Pack Submissions/Dataset/DHI_01_27_22_v1.csv")  
```

##STEP 10: DATAPACK OUTPUT
```{r}  

#WITHDISAGGREGATES  
df_datapack_all<-df_all %>% 
  get_names() %>% 
    dplyr::mutate(`Agency Category` = `fundingagency`)%>%
       dplyr:: mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                             ifelse(`Agency Category` == "USAID/WCF", "USAID",
                                                    ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                                           ifelse(`Agency Category` == "Dedup","Dedupe", "Other"
                                                                 ))))) %>% 
           dplyr::mutate(`Agency Category`= as.character(`Agency Category`))

###Incorporate central america renaming
  df.rename2<- read.csv("C:/Users/vdacosta/Documents/R/COP22 TOOLS/rename.csv" , check.names = FALSE) %>% 
    rename("mech_code"= "﻿mech_code") %>% 
           dplyr::mutate(`mech_code`= as.character(`mech_code`))

 df_datapack_all_join <- left_join(df_datapack_all, df.rename2, by = "mech_code")

df_datapack_all_clean<- df_datapack_all_join %>% 
  dplyr::mutate(countryname2 = case_when(mech_code== "86252" | mech_code== "100129"|  mech_code== "100139"|  mech_code== "81114"|  mech_code== "100110"|  mech_code== "100105"|  mech_code== "84311" |  mech_code== "100130"|  mech_code== "100140"|  mech_code== "81144"|  mech_code== "100106"|  mech_code== "100111"|  mech_code== "84230" |  mech_code== "100131"|  mech_code== "100141"|  mech_code== "81176"|  mech_code== "100112"|  mech_code== "84360" |  mech_code== "81325"|  mech_code== "100113"| mech_code== "100138"|   mech_code== "100143"| mech_code== "100114"|  mech_code== "84730"  ~ Country, TRUE ~ countryname)) %>% 
    select (-c(countryname, Country)) %>% 
    rename("countryname"= "countryname2") 

  
  write_csv(df_datapack_all_clean,("C:/Users/vdacosta/Documents/R/COP22 TOOLS/Datapack_Agency_IM_06_11_22.csv") ) 


#WITHOUTDISAGGREGATES COP21 only
  df_datapack_agg<- df_all %>%
  dplyr::filter(disagg != "KeyPop") %>%
  dplyr::group_by(operatingunit, countryname, fundingagency, mech_code, primepartner, mech_name, indicator, fiscal_year, numeratordenom) %>%
  dplyr::summarise(targets = sum(targets, na.rm = TRUE)) %>% 
  dplyr::ungroup() %>% 
dplyr::mutate(`Agency Category` = `fundingagency`) %>% 
dplyr::mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedup","Dedupe","Other")))) 
    
  write_csv(df_datapack_agg,"C:/Users/vdacosta/Documents/R/COP22 TOOLS/Datapack_Agency_IM_NODISSAGGS_03_24_22.csv")  
  
  
```

##STEP 11: COMMODITIES ONLY Output
```{r}

 df.Commodities_only<- df.Commodities %>% 
    dplyr::filter(`Total Planned Funding` !=0) %>%
     dplyr::mutate(
    `Agency Category` = case_when(
     `Funding Agency` == "USAID/WCF"~ "USAID",
      TRUE ~ `Agency Category` )) %>% 
     dplyr::mutate(`Program Area`= recode (`Program Area`, "c&T"= "C&T")) %>% 
        dplyr::rename("Country" = `Operating Unit`)

 df.Commodities_only <- merge(df.Commodities_only, df.OU_Country, by = "Country")

  write_csv( df.Commodities_only,"COP22 FAST and Data Pack Submissions/Dataset/Commodities_Only_01_27_22_v1.csv")  

```
