---
title: "COP21 FAST Datapack Code"
author: "Jairo Montes & Vanessa Da Costa"
date: "01/26/2022"
output: html_document
---


##Purpose:
This code was developed to integrate real-time notional budget data from the COP22 FAST with COP21 notional target data from  Data Pack and historic expenditure and performance data. 


##Step 1: Installing Packages
R packages are a collection of R functions, complied code and sample data. They are stored under a directory called "library" in the R environment. By default, R installs a set of packages during installation. https://www.tutorialspoint.com/r/r_packages.htm
```{r}
install.packages("tidyr")
install.packages("tidyverse")
install.packages("dplyr")
install.packages("stringr")
install.packages("readxl")
install.packages("here")
#install.packages("googlesheets4")
install.packages("remote")
install.packages("data.table")
install.packages("splitstackshape")
#install.packages("googledrive")
devtools::install_github("USAID-OHA-SI/glamr")
install.packages("purrr")
install.packages("devtools")
remotes::install_github("USAID-OHA-SI/tameDP")
remotes::install_github("USAID-OHA-SI/gophr")
```

##Step 2: Running Libraries
```{r}
library(rmarkdown)
library(stringr)
library(dplyr)
library(tidyverse)
library(tidyr)
library(readr)
library(readxl)
library(here)
library(glamr)
library(purrr)
library(data.table)
#library(googlesheets4)
library(splitstackshape)
#library(googledrive)
library(tameDP)
library(gophr)
library(devtools)
```


##Step 3: Locating the dataset files
The "here()" funcitons constructs paths to your project's files, it will locate the files relative to your project root.
```{r}
  here()
  here("FASTS")
```

##Step 4: Creating functions
Load the following source files to apply the functions necessary to create each data stream
```{R}
source("~/GitHub/EA-Utilities/COP Analytics/test_function.R")
source("~/GitHub/EA-Utilities/COP Analytics/dp_function.R")
source("~/GitHub/EA-Utilities/COP Analytics/esf_function.R")
#source("~/GitHub/EA-Utilities/COP Analytics/OU_Country.csv")
```


5)FAST BY EARMARKS_OU [NOT NEEDED FOR PRIMARY DATASET]
##if code does not work, check to see the column range chosen (AB/Y Denominator should be last column selected). Applies to section 6
```{r}
  #Clean the FAST SCM tab using the function for Earmarks data-roll up to OU level so it can be compared to PLL Earmarks. 
 # FAST_Earmarks_OU <-function(df){
    #nested read_csv. Can be removed and run separately
    df<-read_xlsx(df, "Standard COP Matrix-R", skip=3) %>% 
    dplyr::select('Operating Unit', 'Implementation Year', 'C&T Earmark':'AB/Y Denominator')%>% #adding ab/y numerator and denominator
      dplyr::rename("Fiscal Year" = `Implementation Year`)

    #Convert columns into characters
    df<-df%>%
      dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`))  

    #convert Earmark columns to numeric
    #Change to potential single line code
    df<-df%>%
      dplyr::mutate(`C&T Earmark`=as.numeric(`C&T Earmark`))%>%
      dplyr::mutate(`OVC Earmark`=as.numeric(`OVC Earmark`))%>%
      dplyr::mutate(`Water GHP-State`=as.numeric(`Water GHP-State`))%>%
      dplyr::mutate(`GBV GHP-State`=as.numeric(`GBV GHP-State`)) %>% 
      dplyr::mutate(`AB/Y Numerator`=as.numeric(`AB/Y Numerator`)) %>% 
      dplyr::mutate(`AB/Y Denominator`=as.numeric(`AB/Y Denominator`))

    #remove N/A's
    df <- df %>%drop_na(`Operating Unit`)
    
    #replace NAs with 0s
    df<-df%>%
      mutate_at(vars(`C&T Earmark`:`AB/Y Denominator`),~replace_na(.,0))

    #Group data by OU
    df<- df %>% 
        group_by(`Operating Unit`)  %>%
      summarise(across(where(is.numeric),list( sum)))  %>%
    dplyr::rename("C&T Earmark Fast" = `C&T Earmark_1`)  %>%
    dplyr::rename("OVC Earmark Fast" = `OVC Earmark_1`)  %>%
    dplyr::rename("Water Earmark Fast" = `Water GHP-State_1`)  %>%
    dplyr::rename("GBV Earmark Fast" = `GBV GHP-State_1`) %>% 
    dplyr::rename("AB/Y Numerator" = `AB/Y Numerator_1`) %>% 
    dplyr::rename("AB/Y Denominator" = `AB/Y Denominator_1`) %>%   
    
    return(df)
  }
```  

6)DATAPACK
VDC
```{R}
#load packages
  library(tameDP) #version 3.2.4
  library(purrr)
 
#identify all the Data Pack files (pull in validated datapacks apps.datim.org/datapack)
  files <- list.files("/Users/Vness/Documents/R/COP21 FAST and Data Pack Submissions/DATAPACKS", full.names = TRUE)
  view(files)
#countryname complete,  operating unit blank
df_all<-map_dfr(.x = files, 
                .f= ~ tame_dp(.x, map_names=FALSE)) 

#add in operating unit
df_all<-get_names(df_all)


#tailor Datapack to append to FAST wrangling    
  df_dp<- df_all %>%
    get_names() %>% 
    df_datapack()
  dplyr::filter(disagg != "KeyPop") %>%
  dplyr::group_by(operatingunit, countryname, fundingagency, mech_code, primepartner, mech_name, indicator, fiscal_year, numeratordenom) %>%
  dplyr::summarise(targets = sum(targets, na.rm = TRUE)) %>% 
  dplyr::ungroup() %>% 
  dplyr::rename(`Operating Unit` = operatingunit,
                `Country` = countryname,
                `Funding Agency` = fundingagency,
                `Mechanism ID` = mech_code,
                `Prime Partner Name` = primepartner,
                `Mechanism Name` = mech_name,
                `Fiscal Year` = fiscal_year,
                Indicator = indicator,
                Target = targets) %>%
  dplyr::mutate(`Fiscal Year` = "2022",
                `Data Stream` = "MER",
                `Planning Cycle`="COP21") %>%
dplyr::mutate(`Agency Category` = `Funding Agency`) %>% 
dplyr::mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedupe adjustments Agency","Dedup","Other")))) %>% 
      dplyr::mutate(`Program Area` = dplyr::case_when(Indicator == "HTS_TST_POS" ~ "HTS",
                                                  Indicator == "HTS_TST" ~ "HTS",
                                                  Indicator == "TX_CURR" ~ "C&T",
                                                  Indicator == "TX_NEW" ~ "C&T",
                                                  Indicator == "PrEP_CURR" ~ "PREV",
                                                  Indicator == "PrEP_NEW" ~ "PREV",
                                                  Indicator == "VMMC" ~ "PREV")) 
  
```

##Step 5: READING IN DATASETS AND APPLYING THE FUNCTIONS--> dfs
Read in the FAST & Datapack and clean it using our functions. Pattern is helpful to use here but be careful that you are not picking up more than one dataset.
```{R}
 #Read in Fast files
 # Fast<-list.files(here("COP21 FAST and Data Pack Submissions/FASTS_DRAFTS"),full.names = TRUE)
  Fast<-list.files(here("FASTS"),full.names = TRUE)

  #Data frame for Intervention Budget
  df.Intervention <- purrr::map_dfr(.x = Fast,
                                  .f = ~ FAST_Intervention(.x))

  #Data frame for Cross-Cutting Attribution Budget
  df.CCA <- purrr::map_dfr(.x = Fast,
                         .f = ~ FAST_CCA(.x))

  #Data frame for Initiative Budget
  df.Initiative <- purrr::map_dfr(.x = Fast,
                           .f = ~ FAST_Initiative(.x))

  
  #additional steps for Commodities
  #Data frame for Commodities Budget (prior to merging in identifiers)
  df.Commodities_pre <- purrr::map_dfr(.x = Fast,
                                  .f = ~ FAST_Commodities(.x))

  #Data frame for Mech List
  df.MechsList <- purrr::map_dfr(.x = Fast,
                                       .f = ~ FAST_MECHSLIST(.x))
  
  #Merge in  Mech List to make final Commodities data frame
  #Note that this includes OU but not Operating Unit
  df.Commodities<- left_join(df.Commodities_pre, df.MechsList, by = "Mechanism ID")
  
  #Data frame for Earmarks from FAST by OU
 # df.Earmarks_OU <- purrr::map_dfr(.x = Fast,
                #                .f = ~ FAST_Earmarks_OU(.x))
  
  #Data frame for Earmarks from FAST by IM
  df.Earmarks_IM <- purrr::map_dfr(.x = Fast,
                                 .f = ~ FAST_Earmarks_IM(.x))
 
  #Data frame for Operating Unit and Country
  df.OU_Country <- read.csv (
    "~/GitHub/EA-Utilities/COP Analytics/OU_Country.csv", check.names = FALSE,  
    fileEncoding="UTF-8-BOM")

  #Data frame for Earmarks from PLL
  
 # df.Earmarks_PLL <- read.csv ("COP21 FAST and Data Pack Submissions/Reference/COP21 Earmarks PLL.csv", check.names = FALSE)  
  
  #Merge in Earmarks from PLL dataset to Earmarks from FAST
  #df.Earmarks_OU<- merge(df.Earmarks_PLL,df.Earmarks_OU, by = "Operating Unit") 
  
```
##Step 6: Bind Dataframes
Master COP22 FAST/DATAPACK File
Note: Bind dataframes do not include df. Commodities_pre but only include 'df.Commodities'
```{r}
  #bind all data frames
  df.COP22PlanningDataset<-bind_rows(df.Intervention,df.CCA, df.Initiative,  
                                    df.Commodities, df.Earmarks_IM) %>% 
  COP22_master_clean()

  #add in Operating Unit
   df.COP22PlanningDataset <- merge(df.COP22PlanningDataset, df.OU_Country, by = "Country")
    
  #RUN IF HAVE FASTS AND DATAPACK 
  df.COP22PlanningDataset_Final <-bind_rows(df.COP22PlanningDataset,df_dp) 


```

##Step 7:Review Data
```{r}
  unique(df.COP22PlanningDataset$`Data Stream`)
  unique(df.COP22PlanningDataset_Final$`Program Area`)
  glimpse(df.Earmarks_OU)
  df.COP21PlanningDataset%>%
    group_by(`Fiscal Year`, `Data Stream`) %>%
    filter (`Fiscal Year`==2023)  %>%
    summarize (sum(`Total Planned Funding`))
```

##Step 8: Write the file to output folder.
Update the date
```{r}
  write_csv(df.COP21PlanningDataset_Final,"COP21 FAST and Data Pack Submissions/Dataset/FAST_DATAPACK_05_25_v1.csv")
```

##Additional Steps
##Step 9:FAST by Earmarks Dataset [NOT NEEDED FOR PRIMARY DATASET]
EARMARKS OU DATASET
```{r}  
  write.table(df.Earmarks_OU, paste ('COP21 FAST and Data Pack Submissions/Dataset/COP21_FAST_EARMARKS_OU_', Sys.Date(), ".txt"), sep ="\t", row.names=FALSE)
```
##Step 10 :Aggregate COVID-19 Tab into its own dataset
#Check with Christy to review this part of the code.
```{r}  
  
  FAST_COVID<-function(df){
  #nested read_csv. Can be removed and run separately
  df<-read_xlsx(df, "COVID-E", skip=3)
  
  # Drop columns you don't need and rename  
  df<- df %>% dplyr::rename("Prime Partner Name" = `Partner Name`) 

  #Add in agency category column to group agencies
  df<- df %>% dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
    mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "USAID/WCF", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedup", "Dedup","Other"))))) %>% 
    dplyr::mutate(`Agency Category`= as.character(`Agency Category`))%>% 
    dplyr::mutate(`Target completion date of activity`= as.character(`Target completion date of activity`)) %>% 
    dplyr::mutate(`Mechanism ID`= as.character(`Mechanism ID`)) %>% 
    dplyr::mutate(`Funding Agency`= as.character(`Funding Agency`)) %>% 
    dplyr::mutate(`Mechanism Name`= as.character(`Mechanism Name`)) %>% 
    dplyr::mutate(`Prime Partner Name`= as.character(`Prime Partner Name`)) %>% 
    dplyr::mutate(`Category`= as.character(`Category`)) %>% 
    dplyr::mutate(`Brief description of activity` = as.character(`Brief description of activity`)) %>% 
    dplyr::mutate(`Implementation Timing`= as.character(`Implementation Timing`)) %>% 
    dplyr::mutate(`Priority-rank in order`= as.character(`Priority-rank in order`)) %>% 
    dplyr::mutate(`Are GF or other donors funding any activity in this category?`= as.character(`Are GF or other donors funding any activity in this category?`)) %>% 
    dplyr::mutate(`How has PEPFAR coordinated with GF or other donors to ensure there is not duplication of funding for these activities or sites?`= as.character(`How has PEPFAR coordinated with GF or other donors to ensure there is not duplication of funding for these activities or sites?`)) %>% 
    dplyr::mutate(`Estimate number of patients, HCW and/or sites that would be supported with these activities. (e.g. implement IPC for 100 sites and 400 HCW)`= as.character(`Estimate number of patients, HCW and/or sites that would be supported with these activities. (e.g. implement IPC for 100 sites and 400 HCW)`)) %>% 
    dplyr::mutate(`Target completion date of activity`= as.character(`Target completion date of activity`)) %>% 
    dplyr::mutate(`Approved?`= as.character(`Approved?`)) %>% 
    drop_na("Funding Agency")
    
  return(df)
  }

  df.COVID19 <- purrr::map_dfr(.x = Fast,
                                .f = ~ FAST_COVID(.x))
    
  df.MechsList_short <-df.MechsList %>% 
      dplyr::select('Mechanism ID','Operating Unit') %>% 
      dplyr::rename("Country" = `Operating Unit`) 
    
  df.COVID19_2<- left_join(df.COVID19, df.MechsList_short, by = "Mechanism ID") %>% 
     relocate('Country') %>% 
     dplyr::mutate(`Target completion date of activity`= recode (`Target completion date of activity`, "Q4 FY22"= "2022-09-30")) %>%
     dplyr::mutate(`Target completion date of activity`= recode (`Target completion date of activity`, "Q4 FY21"= "2021-09-30")) 

  
  df.COVID19_Final<- left_join(df.COVID19_2, df.OU_Country, by = "Country") 

  
    write_csv(df.COVID19_Final,"COP21 FAST and Data Pack Submissions/Dataset/COVID19_04_15_v1.csv")
```

##STEP 11: Aggregate DHI Investments Tab into its own dataset
```{r}  
    
df.DHI <-df.Intervention %>% 
      dplyr::rename("Country" = `Operating Unit`) %>% 
      dplyr::select('Country':'Digital Health Investments', 'Data Stream') %>% 
      agency_category_fast()
      distinct(.keep_all= TRUE)
      
      glimpse(df.DHI)
  
   df.DHI_Final <- merge(df.DHI, df.OU_Country, by = "Country")
  
   write_csv(df.DHI_Final,"COP21 FAST and Data Pack Submissions/Dataset/DHI_05_25_v1.csv")  
```

##STEP 12: DATAPACK OUTPUT
```{r}  

#WITHDISAGGREGATES     
df_datapack_all <-df_all %>%  
dplyr::mutate(`Agency Category` = `fundingagency`) %>% 
dplyr::mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedupe adjustments Agency","Dedup","Other")))) 

  write_csv(df_datapack_all,"COP21 FAST and Data Pack Submissions/Dataset/Datapack_Master_05_25_v1.csv")  

#WITHOUTDISAGGREGATES
  df_datapack_agg<- df_all %>%
  dplyr::filter(disagg != "KeyPop") %>%
  dplyr::group_by(operatingunit, countryname, fundingagency, mech_code, primepartner, mech_name, indicator, fiscal_year, numeratordenom) %>%
  dplyr::summarise(targets = sum(targets, na.rm = TRUE)) %>% 
  dplyr::ungroup() %>% 
dplyr::mutate(`Agency Category` = `fundingagency`) %>% 
dplyr::mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedupe adjustments Agency","Dedup","Other")))) 
    
  write_csv(df_datapack_agg,"COP21 FAST and Data Pack Submissions/Dataset/Datapack_Agg_Master_05_25_v1.csv")  
```

##STEP 13: COMMODITIES ONLY Output
```{r}
JMq``````````
#COMMODITIES
 df.Commodities_only<- df.Commodities %>% 
    dplyr::filter(`Total Planned Funding` !=0) %>%
     dplyr::mutate(
    `Agency Category` = case_when(
     `Funding Agency` == "USAID/WCF"~ "USAID",
      TRUE ~ `Agency Category` )) %>% 
     dplyr::mutate(`Program Area`= recode (`Program Area`, "c&T"= "C&T")) %>% 
        dplyr::rename("Country" = `Operating Unit`)

 df.Commodities_only <- merge(df.Commodities_only, df.OU_Country, by = "Country")

  write_csv( df.Commodities_only,"COP21 FAST and Data Pack Submissions/Dataset/Commodities_Only_06_03_v1.csv")  

```
