---
title: "COP21 FAST Datapack Code"
author: "Jairo Montes & Vanessa Da Costa"
date: "05/19/2021"
output: html_document
---


##Purpose:
This code was developed to integrate real-time notional budget data from the COP21 FAST with COP21 notional target data from  Data Pack and historic expenditure and performance data. 


##Step 1: Installing Packages
R packages are a collection of R functions, complied code and sample data. They are stored under a directory called "library" in the R environment. By default, R installs a set of packages during installation. https://www.tutorialspoint.com/r/r_packages.htm
```{r}
install.packages("tidyr")
install.packages("tidyverse")
install.packages("dplyr")
install.packages("stringr")
install.packages("readxl")
install.packages("here")
install.packages("googlesheets4")
install.packages("remote")
install.packages("data.table")
install.packages("splitstackshape")
install.packages("googledrive")
install.packages("devtools")
devtools::install_github("USAID-OHA-SI/glamr")
remotes::install_github("ICPI/ICPIutilities")
install.packages("purrr")
install.packages("devtools")
```

##Step 2: Running Libraries
```{r}
library(rmarkdown)
library(stringr)
library(dplyr)
library(tidyverse)
library(tidyr)
library(readr)
library(readxl)
library(here)
library(ICPIutilities)
library(glamr)
library(purrr)
library(here)
library(data.table)
library(googlesheets4)
library(splitstackshape)
library(googledrive)
library(tameDP)
```

##Step 3: Locating the dataset files
The "here()" funcitons constructs paths to your project's files, it will locate the files relative to your project root.
```{r}
  here()
  here("FASTS")
```

##Step 4: Creating functions
By first creating the functions, it will allow us to, later on, clean up/transform the datasets according to variables that are needed to bind the data.

1) FAST BY INTERVENTION
```{r}
  FAST_Intervention<-function(df){
  #nested read_csv. Can be removed and run separately
  df<-read_xlsx(df, "Standard COP Matrix-R", skip=3)
  
  # Drop columns you don't need and rename  
  df<- df %>% dplyr::select( -c('Global','Prime Partner DUNS','Award Number',
                                'Appropriation Year',  'Initiative',
                                'Funding Category','GAP':'ESF', 
                                'Water':'Digital Health Investments'))%>%
    dplyr::rename("Prime Partner Name" = `Partner Name`,
                  "Is Indigenous Prime Partner" = `Is Indigenous Partner?`,
                  "Prime Partner Type" = `Partner Org Type`,
                  "Fiscal Year" = `Implementation Year`,
                  "Program Area" = `Major Program`,
                  "Sub Program Area" = `Sub Program`,
                  "Interaction Type" = `SD/NSD`,
                  "Beneficiary" = `Major Beneficiary`,
                  "Sub Beneficiary" = `Minor Beneficiary`,
                  "COP Budget New Funding" = `Total New Funding Sources`,
                  "COP Budget Pipeline" = `Applied Pipeline Amount`) %>% 
    
    dplyr::mutate(`Data Stream`="FSD")   
  
  #Convert columns into characters
  df<-df%>%
    dplyr::mutate(`Mechanism ID`=as.character(`Mechanism ID`)) %>% 
     dplyr::mutate(`Mechanism Name`=as.character(`Mechanism Name`)) %>% 
    dplyr::mutate(`Operating Unit`= as.character(`Operating Unit`)) %>% 
   dplyr::mutate(`Prime Partner Name`=as.character(`Prime Partner Name`)) %>% 
    dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`)) %>% 
    dplyr::mutate(`Planning Cycle`=as.character(`Planning Cycle`)) %>% 
    dplyr::mutate(`Interaction Type`=as.character(`Interaction Type`)) %>% 
    dplyr::mutate(`Record Type`=as.character(`Record Type`)) %>% 
    dplyr::mutate(`Program Area`=as.character(`Program Area`)) %>% 
    dplyr::mutate(`Sub Program Area`=as.character(`Sub Program Area`)) %>% 
    dplyr::mutate(`Beneficiary`=as.character(`Beneficiary`)) %>% 
    dplyr::mutate(`Sub Beneficiary`=as.character(`Sub Beneficiary`)) %>% 
    dplyr::mutate(`Funding Agency`=as.character(`Funding Agency`)) %>% 
    dplyr::mutate(`Is Indigenous Prime Partner`= as.character(`Is Indigenous Prime Partner`)) %>% 
    dplyr::mutate(`Prime Partner Type`= as.character(`Prime Partner Type`))
  
  #convert budget columns to numeric
  df<-df%>%
    dplyr::mutate(`COP Budget New Funding`=as.numeric(`COP Budget New Funding`))%>%
    dplyr::mutate(`COP Budget Pipeline`=as.numeric(`COP Budget Pipeline`))%>%
    dplyr::mutate(`Total Planned Funding`=as.numeric(`Total Planned Funding`)) %>% 
  
  #remove N/A's
    #Drop all rows without an OU specified 
    drop_na('Operating Unit')
    
    
  #replace NAs with 0s
  df<-df%>%
    mutate_at(vars(`COP Budget New Funding`:`Total Planned Funding`),~replace_na(.,0))
  
  #Add in agency category column to group agencies
  df<- df %>% dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
    mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedup", "Dedup","Other")))) %>% 
    dplyr::mutate(`Agency Category`= as.character(`Agency Category`))

    #recode values to match naming in Financial Integrated Dataset
  df<- df %>%  dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "SD"= "Service Delivery")) %>%
    dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "NSD"= "Non Service Delivery"))
  
  return(df)
}
```
2) FAST by CROSS-CUTTING
```{r}
  #Clean the FAST SCM tab using the function for Cross-Cutting Attribution + Earmark data.
  FAST_CCA<-function(df){
  #nested read_csv. Can be removed and run separately
  df<-read_xlsx(df, "Standard COP Matrix-R", skip=3)
  
  # Drop columns you don't need and rename  
  df<- df %>% dplyr::select( -c('Global','Prime Partner DUNS','Award Number', 'Cost Type',
                                    'Appropriation Year',  'Initiative',
                                    'Funding Category','GAP':'Total Planned Funding', 
                                    'COVID Adaptation-GHP-State':'Digital Health Investments'))%>%
    dplyr::rename("Prime Partner Name" = `Partner Name`,
                "Is Indigenous Prime Partner" = `Is Indigenous Partner?`,
                "Prime Partner Type" = `Partner Org Type`,
                "Fiscal Year" = `Implementation Year`,
                "Program Area" = `Major Program`,
                "Sub Program Area" = `Sub Program`,
                "Interaction Type" = `SD/NSD`,
                "Beneficiary" = `Major Beneficiary`,
                "Sub Beneficiary" = `Minor Beneficiary`) %>%    

    dplyr::mutate(`Data Stream`="FAST Cross Cutting Attribution")
  
  #remove N/A's
  df <- df %>%drop_na(`Operating Unit`)
  
  #replace NAs with 0s
  df<-df%>%
    mutate_at(vars(`Water`:`Key Populations: SW`),~replace_na(.,0))
  
  
  #Using pivot long to shift CCA vertically
  df <- df %>% pivot_longer(cols = `Water`:`Key Populations: SW`,
                            names_to = "Cross-Cutting Attribution",
                            values_to = "Total Planned Funding")
  
  #Convert columns into characters
  df<-df%>%
   dplyr::mutate(`Mechanism ID`=as.character(`Mechanism ID`)) %>% 
     dplyr::mutate(`Mechanism Name`=as.character(`Mechanism Name`)) %>% 
    dplyr::mutate(`Operating Unit`= as.character(`Operating Unit`)) %>% 
   dplyr::mutate(`Prime Partner Name`=as.character(`Prime Partner Name`)) %>% 
    dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`)) %>% 
    dplyr::mutate(`Planning Cycle`=as.character(`Planning Cycle`)) %>% 
    dplyr::mutate(`Interaction Type`=as.character(`Interaction Type`)) %>% 
    dplyr::mutate(`Record Type`=as.character(`Record Type`)) %>% 
    dplyr::mutate(`Program Area`=as.character(`Program Area`)) %>% 
    dplyr::mutate(`Sub Program Area`=as.character(`Sub Program Area`)) %>% 
    dplyr::mutate(`Beneficiary`=as.character(`Beneficiary`)) %>% 
    dplyr::mutate(`Sub Beneficiary`=as.character(`Sub Beneficiary`)) %>% 
    dplyr::mutate(`Funding Agency`=as.character(`Funding Agency`)) %>% 
    dplyr::mutate(`Is Indigenous Prime Partner`= as.character(`Is Indigenous Prime Partner`)) %>% 
    dplyr::mutate(`Prime Partner Type`= as.character(`Prime Partner Type`))
  
  #Convert CCA budget into numeric
  df<-df%>%
    dplyr::mutate(`Total Planned Funding`=as.numeric(`Total Planned Funding`))
  
  #Add in agency category column to group agencies
  df<- df %>% dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
    mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedup", "Dedup","Other")))) %>% 
      dplyr::mutate(`Agency Category`= as.character(`Agency Category`))

  #recode values to match naming in Financial Integrated Dataset
  df<- df %>%  dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "SD"= "Service Delivery")) %>%
  dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "NSD"= "Non Service Delivery"))

  
  return(df)
}
```
3) FAST BY INITIATIVE
```{r}
  #Clean the FAST SCM tab using the function for Initiative data. 
  #****Note that this is not by funding account since the FAST doesn't have applied pipeline by funding account*****
  FAST_Initiative<-function(df){
  #nested read_csv. Can be removed and run separately
  df<-read_xlsx(df, "Standard COP Matrix-R", skip=3)
  #include columns of interest
  df<- df %>%  dplyr::select('Planning Cycle','Operating Unit':'Partner Name', 'Mechanism Name':'Initiative', 'Implementation Year':'Minor Beneficiary', 'Total New Funding Sources':'Applied Pipeline Amount')  %>%
    #Rename to match BUDGET_ER_MER Dataset
    dplyr::rename("Prime Partner Name" =`Partner Name`, 
                   "Is Indigenous Prime Partner" =`Is Indigenous Partner?`,
                   "Prime Partner Type" =`Partner Org Type`,
                   "Initiative Name" =`Initiative`,
                   "Fiscal Year"= `Implementation Year`, 
                   "Program Area" = `Major Program`,
                   "Sub Program Area" = `Sub Program`,
                   "Interaction Type" = `SD/NSD`,
                   "Beneficiary" =`Major Beneficiary`,
                   "Sub Beneficiary" =`Minor Beneficiary`,
                   "New Funding" =`Total New Funding Sources`,
                   "Applied Pipeline" =`Applied Pipeline Amount`,) %>%
    
    #Pivot COP Budget New Funding & COP Budget Pipeline to 'funding_type' with value as 'Total Planned Funding'
    gather(`Funding Type`,`Total Planned Funding`, `New Funding`:`Applied Pipeline`) %>%
    
    #Pivot GAP, GHP-STATE, GHP-USAID to 'funding_account' with value as 'COP Budget New Funding'
    #  gather(funding_account,`COP Budget New Funding`, `GAP`:`GHP-USAID`) %>%
    
    #Create variable 'Data stream' with Initiative
    dplyr::mutate(`Data Stream`="Initiative")  %>% #consider renaming to specify FAST
    
    #Convert columns into characters
    dplyr::mutate(`Initiative Name`= as.character(`Initiative Name`)) %>% 
     dplyr::mutate(`Funding Category`= as.character(`Funding Category`)) %>% 
    dplyr::mutate(`Mechanism Name`=as.character(`Mechanism Name`)) %>% 
   dplyr::mutate(`Mechanism ID`=as.character(`Mechanism ID`)) %>% 
    dplyr::mutate(`Operating Unit`= as.character(`Operating Unit`)) %>% 
   dplyr::mutate(`Prime Partner Name`=as.character(`Prime Partner Name`)) %>% 
    dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`)) %>% 
    dplyr::mutate(`Planning Cycle`=as.character(`Planning Cycle`)) %>% 
    dplyr::mutate(`Interaction Type`=as.character(`Interaction Type`)) %>% 
    dplyr::mutate(`Record Type`=as.character(`Record Type`)) %>% 
    dplyr::mutate(`Program Area`=as.character(`Program Area`)) %>% 
    dplyr::mutate(`Sub Program Area`=as.character(`Sub Program Area`)) %>% 
    dplyr::mutate(`Beneficiary`=as.character(`Beneficiary`)) %>% 
    dplyr::mutate(`Sub Beneficiary`=as.character(`Sub Beneficiary`)) %>% 
    dplyr::mutate(`Funding Agency`=as.character(`Funding Agency`)) %>% 
    dplyr::mutate(`Is Indigenous Prime Partner`= as.character(`Is Indigenous Prime Partner`)) %>% 
    dplyr::mutate(`Prime Partner Type`= as.character(`Prime Partner Type`)) %>% 
  
    #Convert  budget into numeric
    dplyr::mutate(`Total Planned Funding`=as.numeric(`Total Planned Funding`)) %>% 
   
    #Replace NAs in numeric columns
    dplyr::mutate_at(vars(`Total Planned Funding`),~replace_na(.,0))%>%
    
    
    #Drop all rows without an OU specified 
    drop_na('Operating Unit')   %>%
    
    #recode values for different variables as needed
    dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "SD"= "Service Delivery")) %>%
    dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "NSD"= "Non Service Delivery"))  %>%
  

   #Add in agency category column to group agencies
    dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
    mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedup", "Dedup","Other")))) %>% 
      dplyr::mutate(`Agency Category`= as.character(`Agency Category`))
  return(df)
}
```
4) FAST BY COMMODITIES
```{r}
  #Clean the FAST Commodities tab using the function for Initiative data. 
  FAST_Commodities<-function(df){
  #nested read_csv. Can be removed and run separately
  df<-read_xlsx(df, "Commodities-E", skip=3)
  #include columns of interest
  df<- df %>%  
      dplyr::rename("Specify Other Procurement" =`Specify 'Other' Procurement`) %>% 
  dplyr::select('Mechanism ID': 'Beneficiary', 'Initiative Name', 'Major Category': 'Item', 'Specify Other Procurement','Commodity Quantity', 'Unit Price', 'Unit Cost':'Total Item Budget')  %>%
  dplyr::mutate(`Program Area (Service Delivery Only)`= as.character(`Program Area (Service Delivery Only)`)) %>% 
  dplyr::mutate(`Initiative Name`= as.character(`Initiative Name`)) %>% 
  dplyr::mutate(`Major Category`= as.character(`Major Category`)) %>% 
  dplyr::mutate(`Minor Category`= as.character(`Minor Category`)) %>% 
  dplyr::mutate(`Beneficiary`= as.character(`Beneficiary`)) %>% 
  dplyr::mutate(`Item`= as.character(`Item`)) %>% 
  dplyr::mutate(`Specify Other Procurement`= as.character(`Specify Other Procurement`)) %>% 


    #Remove dashes in Facility-based testing and Community-Based Testing temporarily
    dplyr::mutate(`Program Area (Service Delivery Only)`= recode (`Program Area (Service Delivery Only)`, "HTS: Facility-based testing-SD"= "HTS: Facility based testing-SD")) %>%
    dplyr::mutate(`Program Area (Service Delivery Only)`= recode (`Program Area (Service Delivery Only)`, "HTS: Community-based testing-SD"= "HTS: Community based testing-SD")) %>%
    dplyr::mutate(`Program Area (Service Delivery Only)`= recode (`Program Area (Service Delivery Only)`, "HTS: Facility-based testing-NSD"= "HTS: Facility based testing-NSD")) %>%
    dplyr::mutate(`Program Area (Service Delivery Only)`= recode (`Program Area (Service Delivery Only)`, "HTS: Community-based testing-NSD"= "HTS: Community based testing-NSD"))%>%
    dplyr::rename("Program Area" =`Program Area (Service Delivery Only)`) %>%
    separate(col = "Program Area", into=c("Program Area", "Sub Program Area"), sep=":") %>%
        separate(col = "Sub Program Area", into=c("Sub Program Area", "Interaction Type"), sep="-") %>%

    # cSplit("Program Area", ":" ,, drop=FALSE) %>%
     # separate(col = "Program Area_2", into=c("Sub Program Area", "Interaction Type"), sep="-") %>%
    separate(col = "Beneficiary", into=c("Beneficiary", "Sub Beneficiary"), sep=":") %>%
    
    #Rename to match BUDGET_ER_MER Dataset
   # dplyr::rename("Program Area" =`Program Area_1`,) %>%
    dplyr::rename("Commodity Unit Cost" =`Unit Cost`,) %>%
    dplyr::rename("Commodity Unit Price" =`Unit Price`,) %>%
    dplyr::rename("Total Planned Funding" =`Total Item Budget`,) %>%
    dplyr::rename("Commodity Item" =`Item`,) %>%
    
    
    #Create variables
    dplyr::mutate(`Data Stream`="FAST Commodities")  %>%
    dplyr::mutate(`Fiscal Year`="2022")  %>%
    dplyr::mutate(`Planning Cycle`="COP21")  %>%
    dplyr::mutate(`Record Type`="Implementing Mechanism")%>%
    
    
    #Replace NAs in numeric columns
    dplyr::mutate_at(vars(`Total Planned Funding`),~replace_na(.,0))%>%
    
    
    #Convert  character columns into Characters
    dplyr::mutate(`Mechanism ID`=as.character(`Mechanism ID`)) %>% 
    dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`)) %>% 
    
    #Convert  numeric  columns to numeric
    dplyr::mutate(`Commodity Unit Cost`=as.numeric(`Commodity Unit Cost`)) %>% 
    dplyr::mutate(`Commodity Unit Price`=as.numeric(`Commodity Unit Price`)) %>% 
    dplyr::mutate(`Commodity Quantity`=as.numeric(`Commodity Quantity`)) %>% 
    dplyr::mutate(`Total Planned Funding`=as.numeric(`Total Planned Funding`)) %>% 
    
    #Drop all rows without a MECH ID specified 
    drop_na('Mechanism ID') %>%
    
    #recode values for different variables as needed
    dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "SD"= "Service Delivery")) %>%
    dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "NSD"= "Non Service Delivery"))  %>%
    
    #Add dashes back in Facility-based testing and Community-Based Testing 
    dplyr::mutate(`Sub Program Area`= recode (`Sub Program Area`, "Facility based testing"= "Facility-based testing")) %>%
    dplyr::mutate(`Sub Program Area`= recode (`Sub Program Area`, "Community based testing"= "Community-based testing"))  
  
  return(df)
}
```
5) Mechs-List (Reference List)
```{r}
  #The commodities tab doesn't have certain identifiers (Funding Agency, Mechanism Name, Prime Partner) so this list will be used later to join to make the final commodities data frame
  FAST_MECHSLIST<-function(df){
  #nested read_csv. Can be removed and run separately
  df<-read_xlsx(df, "Mechs List-R")
#include columns of interest
  df<- df %>%  
    dplyr::select ("Operating Unit", "Funding Agency", "Partner Name","Mechanism Name", "Mechanism ID", "Is Indigenous Partner?") %>% 
    #Rename to match BUDGET_ER_MER Dataset
    dplyr::rename( 
                   "Prime Partner Name" =`Partner Name`, 
                   "Is Indigenous Prime Partner" =`Is Indigenous Partner?`,
    ) %>%
  
    dplyr::mutate(`Mechanism ID`=as.character(`Mechanism ID`)) %>% 
    
  #Add in agency category column to group agencies
  dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
    mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedup", "Dedup","Other"))))
  
  return(df)
  } 
```
6A)FAST BY EARMARKS_OU [NOT NEEDED FOR PRIMARY DATASET]
```{r}
  #Clean the FAST SCM tab using the function for Earmarks data-roll up to OU level so it can be compared to PLL Earmarks. 
 # FAST_Earmarks_OU <-function(df){
    #nested read_csv. Can be removed and run separately
    df<-read_xlsx(df, "Standard COP Matrix-R", skip=3) %>% 
    dplyr::select('Operating Unit', 'Implementation Year', 'C&T Earmark':'GBV GHP-State')%>%
      dplyr::rename("Fiscal Year" = `Implementation Year`)

    #Convert columns into characters
    df<-df%>%
      dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`))  

    #convert Earmark columns to numeric
    df<-df%>%
      dplyr::mutate(`C&T Earmark`=as.numeric(`C&T Earmark`))%>%
      dplyr::mutate(`OVC Earmark`=as.numeric(`OVC Earmark`))%>%
      dplyr::mutate(`Water GHP-State`=as.numeric(`Water GHP-State`))%>%
      dplyr::mutate(`GBV GHP-State`=as.numeric(`GBV GHP-State`))

    #remove N/A's
    df <- df %>%drop_na(`Operating Unit`)
    
    #replace NAs with 0s
    df<-df%>%
      mutate_at(vars(`C&T Earmark`:`GBV GHP-State`),~replace_na(.,0))

    #Group data by OU
    df<- df %>% 
        group_by(`Operating Unit`)  %>%
      summarise(across(where(is.numeric),list( sum)))  %>%
    dplyr::rename("C&T Earmark Fast" = `C&T Earmark_1`)  %>%
    dplyr::rename("OVC Earmark Fast" = `OVC Earmark_1`)  %>%
    dplyr::rename("Water Earmark Fast" = `Water GHP-State_1`)  %>%
    dplyr::rename("GBV Earmark Fast" = `GBV GHP-State_1`) 
    
    return(df)
  }
```  
6B)FAST BY EARMARKS_IM
```{r}
  #Clean the FAST SCM tab using the function for Earmarks data-pivot long by IM
  FAST_Earmarks_IM<-function(df){
    #nested read_csv. Can be removed and run separately
    df<-read_xlsx(df, "Standard COP Matrix-R", skip=3)
    
    # Drop columns you don't need and rename  
    df<- df %>% dplyr::select(-c('Global','Prime Partner DUNS','Award Number',
                                  'Appropriation Year',  'Initiative',
                                  'Funding Category','GAP':'Total Planned Funding', 
                                  'Water':'COVID Adaptation-Applied Pipeline'))%>%
      dplyr::rename("Prime Partner Name" = `Partner Name`,
                    "Is Indigenous Prime Partner" = `Is Indigenous Partner?`,
                    "Prime Partner Type" = `Partner Org Type`,
                    "Fiscal Year" = `Implementation Year`,
                    "Program Area" = `Major Program`,
                    "Sub Program Area" = `Sub Program`,
                    "Interaction Type" = `SD/NSD`,
                    "Beneficiary" = `Major Beneficiary`,
                    "Sub Beneficiary" = `Minor Beneficiary`) %>% 
      dplyr::mutate(`Data Stream`="FAST Earmark")   
    
    #remove N/A's
    df <- df %>%drop_na(`Operating Unit`)
    
    #replace NAs with 0s
    df<-df%>%
      mutate_at(vars(`C&T Earmark`:`GBV GHP-State`),~replace_na(.,0))
    
    #Using pivot long to shift Earmarks vertically
    df <- df %>% pivot_longer(cols = `C&T Earmark`:`GBV GHP-State`,
                              names_to = "Earmark",
                              values_to = "Total Planned Funding")
    
    #Convert columns into characters
    df<-df%>%
      dplyr::mutate(`Mechanism ID`=as.character(`Mechanism ID`)) %>% 
     dplyr::mutate(`Mechanism Name`=as.character(`Mechanism Name`)) %>% 
    dplyr::mutate(`Operating Unit`= as.character(`Operating Unit`)) %>% 
   dplyr::mutate(`Prime Partner Name`=as.character(`Prime Partner Name`)) %>% 
    dplyr::mutate(`Fiscal Year`= as.character(`Fiscal Year`)) %>% 
    dplyr::mutate(`Planning Cycle`=as.character(`Planning Cycle`)) %>% 
    dplyr::mutate(`Interaction Type`=as.character(`Interaction Type`)) %>% 
    dplyr::mutate(`Record Type`=as.character(`Record Type`)) %>% 
    dplyr::mutate(`Program Area`=as.character(`Program Area`)) %>% 
    dplyr::mutate(`Sub Program Area`=as.character(`Sub Program Area`)) %>% 
    dplyr::mutate(`Beneficiary`=as.character(`Beneficiary`)) %>% 
    dplyr::mutate(`Sub Beneficiary`=as.character(`Sub Beneficiary`)) %>% 
    dplyr::mutate(`Funding Agency`=as.character(`Funding Agency`)) %>% 
    dplyr::mutate(`Digital Health Investments`=as.character(`Digital Health Investments`)) %>% 
    dplyr::mutate(`Is Indigenous Prime Partner`= as.character(`Is Indigenous Prime Partner`)) %>% 
    dplyr::mutate(`Prime Partner Type`= as.character(`Prime Partner Type`))
    
    #Convert Earmark budget into numeric
    df<-df%>%
      dplyr::mutate(`Total Planned Funding`=as.numeric(`Total Planned Funding`))
    
    #Add in agency category column to group agencies
    df<- df %>% dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
      mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                        ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                               ifelse(`Agency Category` =="Dedup", "Dedup","Other")))) %>% 
        dplyr::mutate(`Agency Category`= as.character(`Agency Category`))
    #recode values to match naming in Budget-ER-MER Dataset
    df<- df %>%  dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "SD"= "Service Delivery")) %>%
      dplyr::mutate(`Interaction Type`= recode (`Interaction Type`, "NSD"= "Non Service Delivery"))
    
    return(df)
  }
```

7)DATAPACK
```{R}
#load packages
  library(tameDP) #version 3.2.4
  library(purrr)
 
#identify all the Data Pack files (pull in validated datapacks apps.datim.org/datapack)
  files <- list.files("/Users/Vness/Documents/R/COP21 FAST and Data Pack Submissions/DATAPACKS", full.names = TRUE)
  view(files)
#countryname complete,  operating unit blank
df_all<-map_dfr(.x = files, 
                .f= ~ tame_dp(.x, map_names=FALSE)) 

#add in operating unit
df_all<-get_names(df_all)


#tailor Datapack to append to FAST wrangling    
  df_dp<- df_all %>%
  dplyr::filter(disagg != "KeyPop") %>%
  dplyr::group_by(operatingunit, countryname, fundingagency, mech_code, primepartner, mech_name, indicator, fiscal_year, numeratordenom) %>%
  dplyr::summarise(targets = sum(targets, na.rm = TRUE)) %>% 
  dplyr::ungroup() %>% 
  dplyr::rename(`Operating Unit` = operatingunit,
                `Country` = countryname,
                `Funding Agency` = fundingagency,
                `Mechanism ID` = mech_code,
                `Prime Partner Name` = primepartner,
                `Mechanism Name` = mech_name,
                `Fiscal Year` = fiscal_year,
                Indicator = indicator,
                Target = targets) %>%
  dplyr::mutate(`Fiscal Year` = "2022",
                `Data Stream` = "MER",
                `Planning Cycle`="COP21") %>%
dplyr::mutate(`Agency Category` = `Funding Agency`) %>% 
dplyr::mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedupe adjustments Agency","Dedup","Other")))) %>% 
      dplyr::mutate(`Program Area` = dplyr::case_when(Indicator == "HTS_TST_POS" ~ "HTS",
                                                  Indicator == "HTS_TST" ~ "HTS",
                                                  Indicator == "TX_CURR" ~ "C&T",
                                                  Indicator == "TX_NEW" ~ "C&T",
                                                  Indicator == "PrEP_CURR" ~ "PREV",
                                                  Indicator == "PrEP_NEW" ~ "PREV",
                                                  Indicator == "VMMC" ~ "PREV")) 
  
```

##Step 5: READING IN DATASETS AND APPLYING THE FUNCTIONS--> dfs
Read in the FAST & Datapack and clean it using our functions. Pattern is helpful to use here but be careful that you are not picking up more than one dataset.
```{R}
 #Read in Fast files
 # Fast<-list.files(here("COP21 FAST and Data Pack Submissions/FASTS_DRAFTS"),full.names = TRUE)
  Fast<-list.files("FASTS",full.names = TRUE)

  glimpse(Fast)
  #Data frame for Intervention Budget
  df.Intervention <- purrr::map_dfr(.x = Fast,
                                  .f = ~ FAST_Intervention(.x))

  #Data frame for Cross-Cutting Attribution Budget
  df.CCA <- purrr::map_dfr(.x = Fast,
                         .f = ~ FAST_CCA(.x))

  #Data frame for Initiative Budget
  df.Initiative <- purrr::map_dfr(.x = Fast,
                           .f = ~ FAST_Initiative(.x))

  
  #additional steps for Commodities
  #Data frame for Commodities Budget (prior to merging in identifiers)
  df.Commodities_pre <- purrr::map_dfr(.x = Fast,
                                  .f = ~ FAST_Commodities(.x))

  #Data frame for Mech List
  df.MechsList <- purrr::map_dfr(.x = Fast,
                                       .f = ~ FAST_MECHSLIST(.x))
  
  #Merge in  Mech List to make final Commodities data frame
  #Note that this includes OU but not Operating Unit
  df.Commodities<- left_join(df.Commodities_pre, df.MechsList, by = "Mechanism ID")
  
  #Data frame for Earmarks from FAST by OU
 # df.Earmarks_OU <- purrr::map_dfr(.x = Fast,
                #                .f = ~ FAST_Earmarks_OU(.x))
  
  #Data frame for Earmarks from FAST by IM
  df.Earmarks_IM <- purrr::map_dfr(.x = Fast,
                                 .f = ~ FAST_Earmarks_IM(.x))
 
  #Data frame for Operating Unit and Country
  df.OU_Country <- read.csv ("COP21 FAST and Data Pack Submissions/Reference/OU_Country.csv" , check.names = FALSE)

  #Data frame for Earmarks from PLL
  
 # df.Earmarks_PLL <- read.csv ("COP21 FAST and Data Pack Submissions/Reference/COP21 Earmarks PLL.csv", check.names = FALSE)  
  
  #Merge in Earmarks from PLL dataset to Earmarks from FAST
  #df.Earmarks_OU<- merge(df.Earmarks_PLL,df.Earmarks_OU, by = "Operating Unit") 
  
```
##Step 6: Bind Dataframes
Master COP21 FAST/DATAPACK File
Note: Bind dataframes do not include df. Commodities_pre but only include 'df.Commodities'
```{r}

 df.COP21PlanningDataset<-bind_rows(df.Intervention,df.CCA, df.Initiative, df.Commodities, df.Earmarks_IM) 
  df.COP21PlanningDataset <- df.COP21PlanningDataset %>%
    dplyr::filter(`Total Planned Funding` !=0) %>%
    dplyr::mutate_at(vars(`COP Budget New Funding`),~replace_na(.,0))%>%
    dplyr::mutate_at(vars(`COP Budget Pipeline`),~replace_na(.,0)) %>% 
    dplyr::mutate(
    `Agency Category` = case_when(
     `Funding Agency` == "USAID/WCF"~ "USAID",
      TRUE ~ `Agency Category` )) %>% 
      dplyr::select('Planning Cycle':'Total Planned Funding','Data Stream', 'Agency Category', 'Cross-Cutting Attribution':'Commodity Unit Cost', 'Earmark') %>% 
     dplyr::mutate(`Program Area`= recode (`Program Area`, "c&T"= "C&T")) %>% 
        dplyr::rename("Country" = `Operating Unit`)

   df.COP21PlanningDataset <- merge(df.COP21PlanningDataset, df.OU_Country, by = "Country")
     
    #RUN IF HAVE FASTS AND DATAPACK 
      df.COP21PlanningDataset_Final <-bind_rows(df.COP21PlanningDataset,df_dp) 


```

##Step 7:Review Data
```{r}
  unique(df.COP21PlanningDataset$`Data Stream`)
  unique(df.COP21PlanningDataset_Final$`Program Area`)
  glimpse(df.Earmarks_OU)
  df.COP21PlanningDataset%>%
    group_by(`Fiscal Year`, `Data Stream`) %>%
    filter (`Fiscal Year`==2022)  %>%
    summarize (sum(`Total Planned Funding`))
```

##Step 8: Write the file to output folder.
Update the date
```{r}
  write_csv(df.COP21PlanningDataset_Final,"COP21 FAST and Data Pack Submissions/Dataset/FAST_DATAPACK_05_17_v1.csv")
```

##Additional Steps
##Step 9:FAST by Earmarks Dataset [NOT NEEDED FOR PRIMARY DATASET]
EARMARKS OU DATASET
```{r}  
  write.table(df.Earmarks_OU, paste ('COP21 FAST and Data Pack Submissions/Dataset/COP21_FAST_EARMARKS_OU_', Sys.Date(), ".txt"), sep ="\t", row.names=FALSE)
```
##Step 10 :Aggregate COVID-19 Tab into its own dataset
```{r}  
  
  FAST_COVID<-function(df){
  #nested read_csv. Can be removed and run separately
  df<-read_xlsx(df, "COVID-E", skip=3)
  
  # Drop columns you don't need and rename  
  df<- df %>% dplyr::rename("Prime Partner Name" = `Partner Name`) 

  #Add in agency category column to group agencies
  df<- df %>% dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
    mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "USAID/WCF", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedup", "Dedup","Other"))))) %>% 
    dplyr::mutate(`Agency Category`= as.character(`Agency Category`))%>% 
    dplyr::mutate(`Target completion date of activity`= as.character(`Target completion date of activity`)) %>% 
    dplyr::mutate(`Mechanism ID`= as.character(`Mechanism ID`)) %>% 
    dplyr::mutate(`Funding Agency`= as.character(`Funding Agency`)) %>% 
    dplyr::mutate(`Mechanism Name`= as.character(`Mechanism Name`)) %>% 
    dplyr::mutate(`Prime Partner Name`= as.character(`Prime Partner Name`)) %>% 
    dplyr::mutate(`Category`= as.character(`Category`)) %>% 
    dplyr::mutate(`Brief description of activity` = as.character(`Brief description of activity`)) %>% 
    dplyr::mutate(`Implementation Timing`= as.character(`Implementation Timing`)) %>% 
    dplyr::mutate(`Priority-rank in order`= as.character(`Priority-rank in order`)) %>% 
    dplyr::mutate(`Are GF or other donors funding any activity in this category?`= as.character(`Are GF or other donors funding any activity in this category?`)) %>% 
    dplyr::mutate(`How has PEPFAR coordinated with GF or other donors to ensure there is not duplication of funding for these activities or sites?`= as.character(`How has PEPFAR coordinated with GF or other donors to ensure there is not duplication of funding for these activities or sites?`)) %>% 
    dplyr::mutate(`Estimate number of patients, HCW and/or sites that would be supported with these activities. (e.g. implement IPC for 100 sites and 400 HCW)`= as.character(`Estimate number of patients, HCW and/or sites that would be supported with these activities. (e.g. implement IPC for 100 sites and 400 HCW)`)) %>% 
    dplyr::mutate(`Target completion date of activity`= as.character(`Target completion date of activity`)) %>% 
    dplyr::mutate(`Approved?`= as.character(`Approved?`)) %>% 
    drop_na("Funding Agency")
    
  return(df)
  }

  df.COVID19 <- purrr::map_dfr(.x = Fast,
                                .f = ~ FAST_COVID(.x))
    
  df.MechsList_short <-df.MechsList %>% 
      dplyr::select('Mechanism ID','Operating Unit') %>% 
      dplyr::rename("Country" = `Operating Unit`) 
    
  df.COVID19_2<- left_join(df.COVID19, df.MechsList_short, by = "Mechanism ID") %>% 
     relocate('Country') %>% 
     dplyr::mutate(`Target completion date of activity`= recode (`Target completion date of activity`, "Q4 FY22"= "2022-09-30")) %>%
     dplyr::mutate(`Target completion date of activity`= recode (`Target completion date of activity`, "Q4 FY21"= "2021-09-30")) 

  
  df.COVID19_Final<- left_join(df.COVID19_2, df.OU_Country, by = "Country") 

  
    write_csv(df.COVID19_Final,"COP21 FAST and Data Pack Submissions/Dataset/COVID19_04_15_v1.csv")
```

##STEP 11: Aggregate DHI Investments Tab into its own dataset
```{r}  
     
df.DHI <-df.Intervention %>% 
      dplyr::rename("Country" = `Operating Unit`) %>% 
      dplyr::select('Country':'Digital Health Investments', 'Data Stream') %>% 
    dplyr::mutate(`Agency Category` = `Funding Agency`)%>%
      mutate(`Agency Category` = ifelse(`Agency Category` == "USAID", "USAID",
                                      ifelse(`Agency Category` == "USAID/WCF", "USAID",
                                      ifelse(`Agency Category` == "HHS/CDC", "CDC",
                                             ifelse(`Agency Category` =="Dedup", "Dedup","Other"))))) %>%
      distinct(.keep_all= TRUE)
  
   df.DHI_Final <- merge(df.DHI, df.OU_Country, by = "Country")
  write_csv(df.DHI_Final,"COP21 FAST and Data Pack Submissions/Dataset/DHI_05_19_v1.csv")  






