---
title: |
  | \LARGE \textbf{\textcolor{usaid_red}{FY`r fiscal_yr` EXPENDITURE REPORT}}
  | \LARGE \textbf{\textcolor{usaid_red}{DATA QUALITY REPORT}}
  | \LARGE \textbf{\textcolor{usaid_red}{`r params$ou`}}
output: pdf_document
header-includes: 
  - \usepackage{xcolor}
  - \usepackage{sectsty}
  - \definecolor{usaid_red}{RGB}{186, 12, 47}
  - \definecolor{usaid_blue}{RGB}{0, 47, 108}
  
  - \allsectionsfont{\Large \raggedright \color{usaid_red}}
  
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{graphicx}
  - \usepackage{multirow}
  
  - \usepackage{fancyhdr}
  
  - \pagestyle{fancy}
  - \fancyhead[L,R,C]{}
  - \fancyhead[C]{\bfseries For USAID Internal Use Only}
  
  - \fancypagestyle{plain}{
      \setlength{\headheight}{59pt}
      \fancyhead[R]{\includegraphics[width=5cm]{Horizontal_RGB_294.png}}
      \renewcommand{\headrulewidth}{0pt}}
  
# https://stackoverflow.com/questions/24421235/changing-font-in-pdf-produced-by-rmarkdown
  - \usepackage{helvet}
  - \renewcommand{\familydefault}{\sfdefault}
  
  - \usepackage{titling}
  - \pretitle{\begin{flushleft}}
  - \posttitle{\end{flushleft}}
params:
  ou: "Central Mechanisms"

---
```{r setup, include=FALSE}
#### WARNING: This RMarkdown must be run via iterate_DQR.R, or it will not run. 
####          This file cannot be run standalone.

knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

```{r r-setup, include=FALSE}
library(kableExtra)
library(glamr)
library(gophr)
library(glue)
library(tidyverse)
library(formattable)


#==============================================================================
# If OU is Central Mechanisms, filter for only central mechanisms. Otherwise,
# filter for the OU in question
if (params$ou == "Central Mechanisms"){
  fsd_ou <- df_fsd[df_fsd$mech_code %in% df_centrals$`Mechanism ID`,]
  msd_ou <- df_msd[df_msd$mech_code %in% df_centrals$`Mechanism ID`,]
  hrh_ou <- df_hrh[df_hrh$mech_code %in% df_centrals$`Mechanism ID`,]
}else {
  fsd_ou <- df_fsd %>% filter(operatingunit == params$ou)
  msd_ou <- df_msd %>% filter(operatingunit == params$ou) 
  hrh_ou <- df_hrh %>% filter(operating_unit == params$ou)
}

# Abridge mechanism names so that they fit in the tables without extensive wrapping
fsd_ou <- fsd_ou %>% mutate(mech_name = lapply(mech_name, abridge, max_len=42))
msd_ou <- msd_ou %>% mutate(mech_name = lapply(mech_name, abridge, max_len=42))
hrh_ou <- hrh_ou %>% mutate(mech_names = lapply(mech_name, abridge, max_len=42))

# Generate dataframe of mechs that have too low or too high budget execution
df_perc_exp_im <- exp_perc_budg_check_im(fsd_ou) %>%
  ungroup()

# Generate dataframe of mechanisms that have exactly matching budget and expenditure
df_budg_exp<- fsd_ou %>%
  check_budg_exp() %>% 
  ungroup()

# dataframe of mechs that have >10% Other:Other
df_other <- fsd_ou %>%
  percent_other_check() %>% 
  ungroup()

# dataframe of mechs that have program management > 40% of expenditure
df_pm <- percent_pm_check(fsd_ou) %>%
  ungroup()

# dataframe of mechs that have ER data but no HR data
df_hrh_check <- hrh_check(fsd_ou, hrh_ou)

# dataframe of mechs with too low/high budget execution at the program level
df_perc_exp_pa<- exp_perc_budg_check_pa(fsd_ou) %>%
  ungroup()

# Calculate how many mechs trigger each check. An alternate "empty" table is printed
#     if there are no mechs triggering a certain check
total_num_mechs <- length(unique(df_fsd$mech_code))
uniq_pe_im <- length(unique(df_perc_exp_im$mech_code))
uniq_pe_pa <- length(unique(df_perc_exp_pa$mech_code))
uniq_budg_exp <- length(unique(df_budg_exp$mech_code))
uniq_other <-length(unique(df_other$mech_code))
uniq_pm <- length(unique(df_pm$mech_code))
uniq_hrh <- length(unique(df_hrh_check$mech_code))

```

# \textcolor{usaid_red}{INTRODUCTION}
Post Q4 pre-clean data release the Expenditure Analysis team ran 5 data checks to ensure data quality of the IP submitted templates. The EA team performed the following checks and compiled results for `r params$ou`.  The errors listed below are NOT required to address and are meant to flag potential issues/concerns for further investigation if and as desired.   The mission can determine any subsequent action based on these findings; however, we recommend reaching out to the partner for any concerning issues to have them correct their template and re-submit to DATIM when it reopens for the cleaning window from December 1-17, 2021.

Data quality checks performed include: 
\begin{itemize}
  \item Percent variance between the FAST budget and reported expenditures
  \item Percent of expenditures categorized with the cost category: sub-cost category “Other: Other”
  \item Percent program management expenditures of the FAST budget
  \item Cross data validation between select MER indicators and expenditures
\end{itemize}

# EXPENDITURE AS A PERCENTAGE OF BUDGET
The following table shows mechanisms that spent less than 80\% or more than 120\% of their budget. Missions may consider following up with partners to ensure 1.) reporting is accurate and there have been no errors, and/or 2.) ensure that they have access to rationale/context for the apparent misalignment between expenditure and budget.  This will be especially important for any IMs with program performance concerns. At the end of this report, is another similar table that performs the same check but for each program area for the mechanism.

```{r, echo=FALSE}
# List of column names to use for the actual printed table
col_perc_exp_im <- c("Mechanism Code", "Mechanism Name", 
                  "Total Expenditure", "Total Budgeted", "% Total Expenditures of Budget")

# If any mechanism triggers this check, apply formatting for currency and percents
if (uniq_pe_im !=0){
  df_pe_im <- df_perc_exp_im %>%
    mutate(sum_fast = currency(sum_fast)) %>%
    mutate(sum_exp = currency(sum_exp)) %>%
    mutate(perc_exp = percent(perc_exp)) %>%
    mutate(perc_exp = as.character(perc_exp))
  # Set nulls to hyphens
  df_pe_im$perc_exp[is.na(df_pe_im$perc_exp)] <- '-'

  # Kable formatting is not intuitive, and more art than science. Order of functions matter, unlike GT.
  # See here: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
  df_pe_im %>%
    # longtable is required for tables that may span across multipel pages
    kbl(format = "latex", col.names = col_perc_exp_im, longtable=T,
        # Set escape to "TRUE" due to formatting above requiring it for LaTeX to read it correctly
        booktabs = T, escape=T, caption = "Expenditure as Percent of Budget, IM-level") %>%
    # repeat_header is helpful for readers, as each new page a table is on will repeat the header
    # hold_position is required, otherwise Kables float to the bottom of the page
    kable_paper(latex_options = c("hold_position", "repeat_header"),
                font_size = 9)%>%
    # One potential workaround to "hiding" a column required for ordering data, but not wanted for dispaly
    # https://tex.stackexchange.com/questions/98955/auto-resize-tabular-row-height
    row_spec(0, extra_latex_after = "[0.7cm]", bold=TRUE) %>%
    column_spec(1, width = '5em') %>%
    column_spec(2, width = '15em') %>%
    column_spec(3, width = '5em') %>%
    collapse_rows(columns = 1:2, valign="top")
} else{
  # Note: If there is no data because the check was not triggered by any mechanism, an 
  #       empty table like this is presented instead
    data.frame(list("No Mechanisms were flagged")) %>% 
    kbl(format = "latex", col.names = NULL)%>%
    kable_paper(latex_options = c("hold_position"))
}




```
# ZERO VARIANCE BETWEEN THE FAST BUDGET AND REPORTED EXPENDITURES
If budgets exactly match expenditures at the intervention level, they are listed below. This is to flag any mechanisms that may be reporting their expenditures so that they are identical to the budgets set in FAST so that there appears to be 0\% variance. While we would like to achieve close alignment between expenditures and budgets is unlikely that they would match exactly.  Teams may want to consider discussing this with partners and ensure reporting represents actual cash expenditures as directed by PEPFAR guidance.   

```{r, echo = FALSE}
# Table title
budg_exp_title<- glue("Mechanisms with Expenditures Matching Budget FY{fiscal_yr}")
# Table column names
col_names <- c("Mech code", "Mechanism Name", "Intervention",
               "Total Planned Budget", "Total Expenditure")

# Check if any mechs triggered check
if (uniq_budg_exp !=0){
  df_budg_exp %>%
    select(-diff) %>%
    # format columns as currency 
    mutate(sum_fast = currency(sum_fast)) %>%
    mutate(sum_exp = currency(sum_exp)) %>%
    kbl(format = "latex", caption = budg_exp_title, col.names = col_names, longtable=T,
      booktabs=T, escape=T) %>%
    # bold the column headers
    row_spec(0,bold=TRUE) %>% 
    kable_paper(latex_options= c("hold_position", "repeat_header")) %>% 
    column_spec(1, width = '3em') %>%
    column_spec(2, width = '15em') %>%
    column_spec(3, width = '7em') %>%
    collapse_rows (columns = 1:2, valign="top")
} else{
  data.frame(list("No Mechanisms were flagged")) %>% 
  kbl(format = "latex", col.names = NULL)%>%
  kable_paper(latex_options = c("hold_position"))
}

```

# PERCENT OF “OTHER: OTHER” EXPENDITURES
If a mechanism has more than 10\% of total expenditures under the cost category: sub-cost category “Other: Other” it is flagged below. These should be reviewed by the mechanism and recategorized if possible. 

```{r, echo=FALSE}
col_other <- c("Mech code", "Mechanism Name",
               "% of Expenditure")
if (uniq_other != 0){
  df_other %>%
  select(-operatingunit, -cost_category, -sub_cost_category) %>%
    # Format column as percent
  mutate(perc_exp = percent(perc_exp)) %>%
  kbl(format = "latex", col.names = col_other, longtable=T, booktabs=T,
      caption = 'Mechanisms  with 10 Percent Expenditures as “Other: Other”') %>%
  kable_paper(latex_options= c("hold_position", "scale_down")) %>%
    # Bold column header
  row_spec(0,bold=TRUE) %>%
  column_spec(1, width = '3em') %>%
  column_spec(2, width = '25em') %>%

  collapse_rows (columns = 1:2, valign="top")
}else{
  data.frame(list("No Mechanisms were flagged")) %>% 
  kbl(format = "latex", col.names = NULL)%>%
  kable_paper(latex_options = c("hold_position"))
}


```
# PERCENT OF PROGRAM MANAGEMENT EXPENDITURES
The following mechanisms have 40\% or more of expenditures reported as program management. It is up to the mission to follow up with any of these mechanisms to ensure accuracy and/or understand *what specifically is supported under PM and how it contributes to programming* as these questions may arise during performance/COP discussions. Mechanisms should accurately report program management, regardless of percent of expenditures. 
```{r, echo=FALSE}
col_pm <- c("Mechanism Code", "Mechanism Name", "% of PM Expenditure of Total")
if (uniq_pm !=0){
  df_pm %>%
  select(-program) %>%
  mutate(perc_exp = percent(perc_exp)) %>%
  # must change col to characters, to change NaN into hyphen
  mutate(perc_exp = as.character(perc_exp)) %>%
  mutate(perc_exp = case_when(perc_exp == 'NaN'~'-',
                              TRUE ~perc_exp)) %>%
  kbl(format = "latex", col.names = col_pm, longtable=T, booktabs=T,
      caption = "Percent of Program Management Expenditure") %>%
  kable_paper(latex_options= c("hold_position", 'repeat_header')) %>%
  row_spec(0,bold=TRUE) %>%
  column_spec(1, width = '5em') %>%
  column_spec(2, width = '20em') %>%
  collapse_rows(columns = 1:2, valign="top")
}else{
  data.frame(list("No Mechanisms were flagged")) %>% 
  kbl(format = "latex", col.names = NULL)%>%
  kable_paper(latex_options = c("hold_position"))
}

```
# MER AND EXPENDITURE CROSS DATA VALIDATIONS
Typically, we expect expenditures to align roughly with reported MER indicators (e.g. TX_CURR would indicate C&T expenditures, PrEP_NEW Prev-PrEP expenditures, etc.).  This is not a hard and fast rule, and the flexibility in the PEPFAR financial reporting guidance allows for a “rolling up” into umbrella interventions that may obscure this alignment.  At HQ, we are pushing for more consistency and specificity of reporting so that we can accurately estimate our contributions to all major priorities. 

The table below flags if an IM meets a certain threshold of MER results that are reported, but with no corresponding expenditures. It is up to the mission to determine whether any updates to the data are needed within the country context and follow up with any of these mechanisms accordingly. 
```{r, echo=FALSE}
col_cdv <- c("Mechanism Code", "Mechanism Name", " Indicator",
             "Results", "Missing Expenditure Category")

# Use pmap to iterate across all the correct filters and string lists for the MER vs. FSD check
df_cdv <- 
  pmap(list(msd_str_lst, msd_cond, fsd_filter_lst, fsd_str_lst), 
       ~cross_data_check(msd_ou, fsd_ou, msd_filter = ..1, 
                         cond = ..2, fsd_filter = ..3, fsd_str = ..4)) %>%
  # pmap returns a list of data.frames, which we bind together
  bind_rows() %>%
  group_by(indicator)

len_cdv <- length(df_cdv$mech_code)

if (len_cdv !=0){
  df_cdv %>%
  ungroup() %>%
  kbl(format = "latex", col.names = col_cdv, longtable = T, 
      booktabs = T, caption="Mechanisms Flagged for Reporting MER Results but not Related Expenditures") %>%
  kable_paper(latex_options = c("hold_position", "repeat_header",
                                "bordered")) %>%
    # Manually remove last row with extra asterisks: https://github.com/haozhu233/kableExtra/issues/299
  row_spec(1:len_cdv, hline_after=T, extra_latex_after="%") %>%
  row_spec(0,bold=TRUE) %>%
  column_spec(1, width = '7em') %>%
  column_spec(2, width = '15em') %>%
    # Pack_rows is used to organize rows in groups by indicator
  pack_rows(index = setNames(rle(df_cdv$indicator)[[1]],
                             rle(df_cdv$indicator)[[2]]),
            color = "white",
            background = "gray")

} else{
  data.frame(list("No Mechanisms were flagged")) %>% 
    kbl(format = "latex", col.names = NULL)%>%
    kable_paper(latex_options = c("hold_position"))
}
```
# ENSURING ALL MECHANISMS WHICH REPORTED IN ER ALSO REPORT TO THE NEW HRH INVENTORY
The following table shows which mechanisms have reported in Expenditure Reporting but did not report to the HRH Inventory. **All mechanisms which report to ER should also report to the HRH inventory this year.**
Have these highlighted mechanisms complete the HRH template and upload their data to the HRH processor app in DATIM during data cleaning (Dec. 1st - 17th).
```{r, echo=FALSE}
# See first section for explanation on Kable, col_names, etc.
hrh_col_names <- c("Mechanism Code", "Mechanism Name", "Sum Expenditure")
hrh_title <- "Mechanisms that reported expenditure but not HRH data"
if (uniq_hrh != 0){
  df_hrh_check %>%
    mutate(sum_exp = currency(sum_exp)) %>%
    kbl(format = "latex",  col.names = hrh_col_names, longtable=T,booktabs = T, caption = hrh_title) %>%
    kable_paper(latex_options= c("hold_position", 'repeat_header')) %>%
    row_spec(0,bold=TRUE) %>%
    column_spec(2, width = '25em') %>%
    collapse_rows(columns = 1:2, valign="top")

} else {
    data.frame(list("No Mechanisms were flagged")) %>% 
    kbl(format = "latex", col.names = NULL)%>%
    kable_paper(latex_options = c("hold_position"))
}
```
# EXPENDITURE AS A PERCENTAGE OF BUDGET, MECHANISM BY PROGRAM AREA
The following table shows each program area for a mechanism that spent less than 80\% or more than 120\% of their budget. Missions may consider following up with partners to ensure 1.) reporting is accurate and there have been no errors, and/or 2.) ensure that they have access to rationale/context for the apparent misalignment between expenditure and budget.  This will be especially important for any IMs with program performance concerns. 

```{r, echo=FALSE}
# See first section for explanation on Kable, col_names, etc.
col_perc_exp_pa <- c("Mechanism Code", "Mechanism Name", "Program Area",
                  "Total Expenditure", "Total Budgeted", "% of Budget")

if (uniq_pe_pa != 0){
  # Format columns as currency and  percent
  df_pe_pa <- df_perc_exp_pa %>%
  mutate(sum_fast = currency(sum_fast)) %>%
  mutate(sum_exp = currency(sum_exp)) %>%
  mutate(perc_exp = percent(perc_exp)) %>%
  mutate(perc_exp = as.character(perc_exp))
df_pe_pa$perc_exp[is.na(df_pe_pa$perc_exp)] <- '-'

df_pe_pa %>%
  kbl(format = "latex", col.names = col_perc_exp_pa, longtable=T,
      booktabs = T, escape=T, caption = "Expenditure as Percent of Budget, Program Area-level") %>%
  kable_paper(latex_options = c("hold_position", "repeat_header"),
              font_size = 9)%>%
  # https://tex.stackexchange.com/questions/98955/auto-resize-tabular-row-height
  row_spec(0, extra_latex_after = "[1 cm]", bold=T) %>%
  column_spec(1, width = '5em') %>%
  column_spec(2, width = '15em') %>%
  column_spec(3, width = '5em') %>%
  column_spec(6, width='5em') %>%
  # setting latex_hline = "major" fixes behaviours with hlines for first row of a 2nd or 3rd page
  # of a long table, especially when the collapsed row is cut in between two pages
  collapse_rows(columns = 1:2, valign="top", latex_hline= "major")
} else{
    data.frame(list("No Mechanisms were flagged")) %>% 
    kbl(format = "latex", col.names = NULL)%>%
    kable_paper(latex_options = c("hold_position"))
}
```